Illusion Game Validation (Study 1)

### 2022-08-03 study1.Rmd
script

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3,
        mc.cores = 2)

@@

Behavioural findings regarding the Illusion Game.

@@

outliers <- c(
  # Half of the trials are of very short RT
  # Prolific Status: REJECTED
  "60684f29dbfe1bb2059e5e27_rkqoy",
  # Error rate of 47.9%
  # Prolific Status: RETURN REQUESTED
  "61280140171ec546e87ed8cb_qdlgy",
  # Error rate of 46.2%
  # Prolific Status: RETURN REQUESTED
  "614f36fd81c78b7a125c4262_6ax4g",
  # Error rate of 42.1% and very large RT SD
  # Prolific Status: RETURN REQUESTED
  "5d398380b37ab1000111fac3_2nxxh",
  # Block n2 with very short RTs
  # Prolific Status: RETURNED
  "5e860198a846e30497df5189_6e43s",
  # Error rate of 46.2% and short RTs
  # Prolific Status: RETURN REQUESTED
  "615f319bb341cf2f306d858d_qsaq5"
)
```

@@

@@

  if(side == "leftright") {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if(x == "Left") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
    } else {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
    }
  } else {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if(x == "Up") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
    } else {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
    }
    data$Answer <- fct_rev(data$Answer)
  }

@@

```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
best_models <- function(data) {
  models <- list()
  for(i in 1:2) {
    for(j in 1:2) {
      for(k1 in c("", "_log")) {  # c("", "_log")
        for(k2 in c("", "_log")) {  # c("", "_log")
          for(side in c("", "-side")) {
            name <- paste0("dif", k1, i, "-", "str", k2, j, side)
            print(name)
            f <- paste0("poly(Illusion_Difference", 
                        k1,
                        ", ",
                        i,
                        ") * poly(Illusion_Strength",
                        k2, 
                        ", ",
                        j, 
                        ") + (1|Participant)")

@@

model <- brms::brm(formula,
  data = data,
  algorithm = "sampling",
  refresh = 200
)
```

@@

  formula <- brms::bf(
    Error ~ t2(Illusion_Difference, Illusion_Strength, bs = "cr", k=4) + 
      ((1 + Illusion_Difference + Illusion_Strength) | Participant),
    family = "bernoulli"
  )

  model <- brms::brm(formula,
    data = data,
    algorithm = "meanfield",
    refresh = 0
  )

@@

range(data$Illusion_Difference)
range(data$Illusion_Strength)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  algorithm = "meanfield",
  refresh = 0
)
```

@@

@@

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  algorithm = "meanfield",
  refresh = 0
)
```

@@

@@

```{r message=FALSE, warning=FALSE, cache=cache}
# model <- glmmTMB::glmmTMB(
#   Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
#     ((1 + Illusion_Difference + Illusion_Strength) | Participant),
#   family = "binomial",
#   data=data
# )

formula <- brms::bf(
  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  algorithm = "meanfield",
  refresh = 0
)
```

@@

@@

@@

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ Illusion_Side * Illusion_Difference * poly(Illusion_Strength, 2) + 
    ((1 + Illusion_Side + Illusion_Difference + Illusion_Strength) | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  algorithm = "meanfield",
  refresh = 0
)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  algorithm = "meanfield",
  refresh = 0
)
```

>>>>

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3,
        mc.cores = 4,
        brms.algorithm = "sampling",
        brms.backend = "cmdstanr")

@@

The goal of study 1 was to...

@@

outliers <- c(
  # Half of the trials have very short RT
  # Prolific Status: REJECTED
  "60684f29dbfe1bb2059e5e27_rkqoy",
  # Block n2 with very short RTs
  # Prolific Status: RETURNED
  "5e860198a846e30497df5189_6e43s",
  # Error rate of 46.2% and short RTs
  # Prolific Status: RETURNED
  "615f319bb341cf2f306d858d_qsaq5",
  # Error rate of 46.2%
  # Prolific Status: RETURNED
  "614f36fd81c78b7a125c4262_6ax4g",
  # Error rate of 47.9%
  # Prolific Status: RETURN REQUESTED
  "61280140171ec546e87ed8cb_qdlgy",
  # Error rate of 42.1% and very large RT SD
  # Prolific Status: RETURN REQUESTED
  "5d398380b37ab1000111fac3_2nxxh"
)
```

@@

## Data Analysis

The analysis of study focused on the probability of errors as the main outcome variable. For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some insight on the underlying generative model. Next, we tested the performance of various models differing in their specifications, such as: with or without a log-transform on the task difficulty, with or without a 2nd order polynomial term of the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor. We then fitted the best performing model under a Bayesian model, and compared its predicted links with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships at the expense of model simplicity. Finally, we used these models to estimate different values of task difficulty that would lead to similar error rates (2.5, 5 and 25%) when the illusion strength is null.

@@

  if(side == "leftright") {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if(x == "Left") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
    } else if(x == "Right") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
    }
  } else {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if(x == "Up") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
    } else if(x == "Down") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
    }
    data$Answer <- fct_rev(data$Answer)
  }

@@

```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
best_models <- function(data) {
  models <- list()
  for(i in 1:1) {
    for(j in 1:2) {
      for(k1 in c("", "_log")) { 
        for(k2 in c("")) {  
          for(side in c("", "-side")) {
            name <- paste0("dif", k1, i, "-", "str", k2, j, side)
            print(name)
            f <- paste0("poly(Illusion_Difference", 
                        k1,
                        ", ",
                        i,
                        ") * poly(Illusion_Strength",
                        k2, 
                        ", ",
                        j, 
                        ") + (1|Participant)")

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
    (1 | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  algorithm = "sampling",
  refresh = 0
)

@@

  formula <- brms::bf(
    Error ~ t2(Illusion_Difference, Illusion_Strength, bs = "cr", k=4) + 
      (1 | Participant),
    family = "bernoulli"
  )

  model <- brms::brm(formula,
    data = data,
    refresh = 0
  )

@@

# range(data$Illusion_Difference)
# range(data$Illusion_Strength)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
    (1 | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  refresh = 0
)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
# range(data$Illusion_Difference)
# range(data$Illusion_Strength)

@@

data <- filter(data, abs(Illusion_Strength) < 15)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ log(Illusion_Difference) * Illusion_Strength + 
    (1 | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  refresh = 0
)

@@

data <- filter(data, abs(Illusion_Strength) < 90)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ log(Illusion_Difference) * Illusion_Strength + 
    (1 | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  refresh = 0
)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
# range(data$Illusion_Difference)
# range(data$Illusion_Strength)
# unique(data$Illusion_Strength)

@@

```{r message=FALSE, warning=FALSE, cache=cache}
best_models(data)
```

@@

plot_descriptive(data, side = "updown")

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
    (1 | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  refresh = 0
)

@@

model <- brms::brm(formula,
  data = data,
  refresh = 0
)
```
